### 2014-10-24 abstract.tex
Initial commit.

Declarative programming is good for parallel programming but makes it hard for
the programmer to fine tune scheduling and data layout of the program.
some programs require more information in order to execute better.
with our approach we permit the programmer to write declarative programs
but then optimize them later using sensing facts and action facts that
understand and change how the programs are executed.

Linear Meld is a concurrent forward-chaining linear logic programming
language where logical facts can be asserted and retracted in a
structured way. In Linear Meld, a program is seen as a database of
logical facts and a set of derivation rules. The database of facts is
partitioned by the nodes of a graph structure which leads to