### 2014-10-31 intro.tex

Writing parallel programs in sequential languages is hard because manipulating
shared state using multiple threads may result in data race
conditions. Such issues are handled with low level constructs such as locks and
condition variables, requiring a fair amount of effort to get right.
Declarative programming has been hailed as a solution to this issue, since the
problem of implementing the details of parallelism is moved away from the programmer to the
compiler and runtime of the language. The programmer writes code
without having to deal with parallel programming constructs and the compiler
automatically parallelizes the program in order to take advantage of
multiple threads of execution.
The programming paradigm has been adopted with huge success in domain specific
languages such as SQL and MapReduce~\cite{Dean:2008:MSD:1327452.1327492}.
Although more general declarative languages have yet to be as successful, the
future looks promising for this particular approach.

We argue, however, that declarative programming leaves little to no programmer control
over how execution is scheduled or how data is laid out. This introduces some
performance issues because even if the runtime system is able to reasonably
parallelize the program using a general algorithm, there is a lack of specific
information about the program that a compiler does not easily understand. In
turn, such information would make execution better in terms of run time time,
memory usage and scalability.

In this paper, we extend Linear Meld (LM), a declarative language by
\cite{cruz-iclp14,cruz-ppdp14}, with coordination facts that give programmer control over
scheduling and data placement. LM is a linear logic programming language
intended for programs over graph structures and with
support for structured manipulation of mutable state using 
linear logic~\cite{girard-87}. In LM, the nodes of the graph perform computation
through derivation of rules and communicate with other nodes when logical facts
are derived.

The operational semantics of LM allow for plenty of
non-determinism since nodes execute independently and can be executed with
any specific order. Since LM uses mutable state, the order in which nodes are
executed may be crucial in terms of performance. We extended the language with
coordination facts that can be used as a regular fact in the program logic to
improve program execution.
The first kind of coordination facts are
called \emph{sensing facts} and are used to sense information about data
placement (where is the node being executed). This allows the programmer to
write logical rules that depend on the current state of the program. The second
kind of coordination facts are \emph{action facts} that, when deriving, will be
consumed to apply a scheduling operation during execution. This allows the
programmer to prioritize node computation or place nodes in different threads.
Both kinds of facts are \emph{linear facts} since they are consumed and/or change.

To the best of our knowledge, this is the first time that a declarative language allows 
control over execution without a significant change in the language itself and
without using non-declarative operations. This proves specially important since
programs can proven to be correct even in the presence of coordination facts.
Our other two main contributions are as follows: a novel way to coordinate
declarative programs through user control that introduce measurable performance
improvements; several application where we add programmer control to declarative
programs in order to make them perform better.