### 2014-10-31 language.tex

Linear Meld (LM) is a forward-chaining linear logic programming
language~\cite{cruz-iclp14} with roots on Datalog~\cite{Ramakrishnan93asurvey}.
Programs are written as a set of rules of the form
\texttt{a(X), b(Y) -o c(Z)} and are read as following: if \texttt{a(X), b(Y)} (the body of
the rule) is satisfied using some facts \texttt{a(X)} and \texttt{b(Y)} then
\texttt{c(Z)} (the head of the rule) is true.
The facts used during rule derivation are placed in a \emph{database of facts}.
Logical facts are composed of a predicate and a tuple of concrete values,
representing the arguments.
Since LM uses linear logic as its foundation, rule derivations
may delete facts used in the body of the rule. Program execution starts by
adding the \emph{axioms} (the initial facts) of the program to the database.
Next, the rules are recursively applied and the database is updated with new and
deleted facts. When nore more rules are applicable, the program terminates.

LM has been designed for writing graph-based programs. LM accomplishes this by
partitioning the logical facts by the first argument (the \emph{node}) and by
restricting the body of every rule to only refer to the same first argument.
This allows rules to be executed using only facts of a single node. Although the
body of rules is restricted, the head of the rule may refer to any other node as
long as the node variable is refered somewhere in the body of the rule. This allows
derivations of facts in other nodes. The whole database is then composed of smaller
databases (the nodes) and rule derivations allows smaller databases to derive
facts in other databases, creating edges in the graph.

In order to fully understand how the language works, we present a very simple
algorithm: the single source shortest path program~(SSSP). Later in the paper, we
add coordination facts to improve the execution of the program.

The SSSP program in Fig.~\ref{code:shortest_path_program} starts (lines 1-3)
with the declaration of the predicates. Predicates specify the kinds of facts
used in the program. The first predicate, \texttt{edge}, specifies edges between
nodes of the graph, where the third argument represents the weight of the edge.
The \texttt{route} modifier informs the compiler about the structure of the
program graph. The predicates \texttt{shortest} and \texttt{relax} are specified
as linear facts and thus are deleted when used in rules.
The idea of the algorithm is to compute the shortest distance from node
\texttt{@1} to all other nodes in the graph. Every node has a \texttt{shortest}
fact that is improved with new \texttt{relax} facts.
Lines 5-9 declare the axioms of the program: \texttt{edge} facts describe the
graph; \texttt{shortest(A, +00, [])} is the initial shortest distance (infinity)
for all nodes; and \texttt{relax(@1, 0, [@1])} starts the algorithm with the
initial distance from \texttt{@1} to \texttt{@1}.

\begin{figure}[h!]
\scriptsize\begin{Verbatim}[numbers=left]
type route edge(node, node, int).
type linear shortest(node, int, list int).
type linear relax(node, int, list int).

!edge(@1, @2, 3). !edge(@1, @3, 1).
!edge(@3, @2, 1). !edge(@3, @4, 5).
!edge(@2, @4, 1).
shortest(A, +00, []).
relax(@1, 0, [@1]).

relax(A, D1, P1), shortest(A, D2, P2),
D1 < D2
   -o shortest(A, D1, P1),
      {B, W | !edge(A, B, W) |
         relax(B, D1 + W, P1 ++ [B])}.

relax(A, D1, P1), shortest(A, D2, P2),
D1 >= D2
   -o path(A, D2, P2).
\end{Verbatim}
  \caption{Single Source Shortest Path program code.}
  \label{code:shortest_path_program}
\end{figure}
\normalsize

The first rule of the program (lines 11-15) replaces the current path in
\texttt{shortest} with a shorter one in \texttt{relax}. The rule deletes both
\texttt{relax} and \texttt{shortest} facts and derives a new \texttt{shortest}
fact. In lines 14-15, we have a \emph{comprehension} where the program iterates
over the edges of node \texttt{A} and derives a new \texttt{relax} fact at
\texttt{B} with the new shorter distance plus the weight of the edge. For
instance, in Fig.~\ref{fig:shortest_path_program}~(a) we apply rule 1 in node
\texttt{@1} where two new \texttt{relax} facts are derived at node \texttt{@2}
and \texttt{@3}. Fig.~\ref{fig:shortest_path_program}~(b) is the result after
applying the same rule but at node \texttt{2}.

\begin{figure*}[ht]
\begin{center}
  \subfloat[]{\includegraphics[width=0.3\textwidth]{figures/shortest2}}
  \subfloat[]{\includegraphics[width=0.3\textwidth]{figures/shortest3}}
  \subfloat[]{\includegraphics[width=0.3\textwidth]{figures/shortest8}}
\end{center}
\caption{Graphical representation of the SSSP program. Figure (a) represents the
   program after propagating initial distance at node \texttt{@1}, followed by
   Figure (b) where the first rule is applied in node \texttt{@3}. Figure (c)
   represents the state of the final program, where all the shortest paths
   have been computed.}
\label{fig:shortest_path_program}
\end{figure*}

The second rule of the program (lines 17-19) deals with the case where the new
distance in \texttt{relax} is not better than the current one, so it is thrown
away and the current distance is kept.

There many opportunities for concurrency in the SSSP program. For instance,
after applying rule 1 in Fig.~\ref{fig:shortest_path_program}~(a), it is
possible to either apply rules in either node \texttt{@2} or node
\texttt{@3}. In our implementation, we partition the graph into subgraphs that
are processed by multiple threads of execution. Eventually, it is no longer to
possible to apply rules and the final result present in
Fig.~\ref{fig:shortest_path_program}~(c) is achieved.

%On the other hand, LM has no natural matching of data and computation to workers (processes, threads),
%since nodes are a program abstraction and part of the program's logic.
%We view the set of nodes as a graph data structure where workers will perform work.
%A worker is able to process any node, although a node cannot be computed by more than one worker
%at the same time. This disallows the manipulation of a node by multiple workers.