### 2014-11-02 application.tex
Splash BP updates.

To better understand how coordination facts are used, we next present some programs that
take advantage of them.

\subsection{Single Source Shortest Path}
\input{sssp}

\subsection{MiniMax}

\input{minimax}

\subsection{Heat Transfer}

\input{heat-transfer}

\subsection{Splash Belief Propagation}

\input{splash-bp}

\subsection{N Queens}