### 2014-11-04 experiments.tex
Add heat transfer. Add more results.

\subsection{Absolute runtime}

In the paper~\cite{cruz-ppdp14}, the LM virtual machine was measured and
compared against implementations of similar programs but written in other
languages. For instance, the LBP program was found to be around 2 times slower
than GraphLab, while N Queens program was found to be around 10 to 15 slower
than a sequential C implementation. Some of those programs were compared against
a Python implementations and LM faired fairly well. In this section, we measure
the impact of the coordination mechanisms we have implemented for the new
virtual machine.

Fig.~\ref{results:comparison1}

\begin{figure}[h!]
   \begin{center}
      \subfloat[]{\includegraphics[width=9cm]{results/comparison1.png}}
   \end{center}
   \caption{Measuring the overhead of coordination mechanisms. The \textbf{Old}
   bars represent the performance of the original virtual machine, while
   \textbf{New} is the relative performance of the new version that includes
   coordination mechanisms. Most programs show little to no degradation in
   performance.}
   \label{results:comparison1}
\end{figure}