### 2014-11-05 abstract.tex
More corrections.

Declarative programming has been hailed as a promising approach to parallel
programming since it hides the implementation details of parallelism away from the
programmer. However, some control is lost when code is
automatically parallelized, leaving the programmer with very little
opportunities for fine tuning programs.
We extended Linear Meld, a concurrent forward-chaining linear logic programming
language, with coordination facts that allow the programmer to change
computation scheduling and data layout. Our approach allows the programmer to
write declarative parallel programs and then optionally use coordination facts
to fine tune programs, while keeping the program fully declarative and amenable
to correctness proofs. We show several application examples, experimental
results and some advantages of our novel approach.