### 2014-11-06 conclusions.tex
More corrections.

We have presented a way of adding coordination to a declarative language without
change the nature of the language or introducing non-declarative constructs. We
took advantage of the fact that LM uses linear logic, which allows us to derive
coordination facts to delete facts in order to perform actions in the underlying
runtime system.

To the best of our knowledge, this is the first time that a declarative language allows 
control over execution without a significant change in the language itself and
without using non-declarative operations. This is specially important since
programs can be proven to be correct even in the presence of coordination facts.
In order to prove our thesis, we first give a brief overview of the
We think that our particular approach is applicable in distributed systems as we
have seen. Explore functional programming and coordination?