### 2014-11-08 intro.tex

Writing parallel programs in sequential languages is hard because manipulating
shared state using multiple threads may result in data race
conditions. Such issues are handled with low level constructs such as locks and
condition variables, requiring a fair amount of effort to get right.
Declarative programming has been hailed as a solution to this issue, since the
problem of implementing the details of parallelism is moved from the programmer to the
compiler and runtime. The programmer writes code
without having to deal with parallel programming constructs and the compiler
automatically parallelizes the program in order to take advantage of
multiple threads of execution.
The programming paradigm has been adopted with huge success in domain specific
languages such as SQL and MapReduce~\cite{Dean:2008:MSD:1327452.1327492}.
Although general declarative languages have yet to be as successful, the
future looks promising for this particular approach.

The problem with declarative programming is that it leaves little to no programmer control
over how execution is scheduled or how data is laid out, making it hard to achieve
necessary efficiency. This introduces 
performance issues because even if the runtime system is able to reasonably
parallelize the program using a general algorithm, there is a lack of specific
information about the program that a compiler cannot easily deduce. Such
information could make execution better in terms of run time time,
memory usage, and scalability.

In this paper, we introduce Coordinated Linear Meld (CLM), a declarative
language that extends the Linear Meld~(LM) language~\cite{cruz-iclp14,cruz-ppdp14} with coordination facts that give programmer
control over scheduling and data placement. LM is a linear logic programming
language designed for programs that operate on graphs.  The use of linear
logic~\cite{girard-87} supports structured manipulation of mutable
state. In LM, computation is divided so that each node of the graph computes
independently and is allowed to \scare{communicate} with other nodes.
Both computation and communication happen through the derivation of logical
rules (which make up the program).

The CLM language features coordination
primitives that can be used in the same way as any other primitive,
in other words, they are specificied with the same syntax and semantics as the
rest of the programming language. These coordination primitives can be used to
improve program execution based on the state of the program and the underlying
machine. The first kind of coordination facts are called \emph{sensing facts}
and are used to sense information about the system the program is running on,
e.g., scheduling and node placement on threads. The second kind of coordination facts are
\emph{action facts} that when deleted in rules are used to apply a scheduling
operation during execution. Coordination facts allow the programmer to write
logical rules that depend on the current state of the program and then prioritize node
computation or place nodes in different threads.

To the best of our knowledge, this is the first time that a declarative language allows 
control over execution while staying declarative and without resorting to
meta-language constructs. This is crucial to our goal of being able to prove
programs correct since
proofs can be constructed even in the presence of coordination facts.
After briefly discussing related work, we present an overview of the base
language, with an example. In Section~\ref{sec:coordination}, we introduce the
current set of coordination facts,
followed by a description of the changes required to implement the desired
coordination mechanisms. In
Section~\ref{sec:applications} we present several applications and show
coordination can improve programs without destroying clarity or provability.