### 2014-11-08 application.tex

To better understand how coordination facts are used, we next present some programs that
take advantage of them. In our experimental setup, we used a machine with
four AMD Six-Core Opteron TM 8425 HE (2100 MHz) chips (24 cores) and 64 GB of
DDR-2 667MHz (16x4GB) RAM, running GNU/Linux (kernel 3.15.10-201 64 bits).
We compiled our virtual machine using GCC 4.8.3 (g++) with the flags
\texttt{-O3 -std=c++11 -fno-rtti -march=x86-64}~\footnote{Implementation and
   example programs available in \url{http://github.com/.../meld}}.

\subsection{Single Source Shortest Path}
\input{sssp}

\subsection{MiniMax}
\input{minimax}

\subsection{Heat Transfer}
\input{heat-transfer}

\subsection{Splash Belief Propagation}
\input{splash-bp}

\subsection{N Queens}
\input{nqueens}