### 2014-11-08 heat-transfer.tex
Further updates.

In the Heat Transfer algorithm~(HT), we have a graph where heat values are
exchanged between nodes. XXX rewrite The program stops when the new heat values of every node are
$\delta = |H_i - H_{i-1}| \le \epsilon$. The algorithm works
asynchronously since heat values can be updated by using new partial information
coming from neighbor nodes. This increases parallelism since nodes do not need
to synchronize between iterations.

Fig.~\ref{code:ht} shows the HT rules that send new heat values to
neighbor nodes. In the first rule we added \texttt{add-priority} to increase the priority of the neighbor
nodes if the current node has a large $\delta$. The idea is to prioritize the
computation of heat values of nodes (using \texttt{update}) that have a neighbor
that changed significantly. Multiple \texttt{add-priority} facts will
increase the priority of a node so that nodes with multiple large deltas will
have more priority.

\begin{topfig}
\scriptsize\begin{Verbatim}[numbers=left,commandchars=\\\[\]]
new-heat(A, New, Old),
Delta = fabs(New - Old),
Delta > epsilon
   -o {B, W | !edge(A, B, W) |
         new-neighbor-heat(B, A, New),
         update(B), \underline[add-priority(B, Delta)]}.

new-heat(A, New, Old)
fabs(New - Old) <= epsilon
   -o {B, W | !edge(A, B, W) |
         new-neighbor-heat(B, A, New)}.
\end{Verbatim}
  \scap{code:ht}{Coordination code for the Heat Transfer program. We updated the first rule
     to increase the priority of neighbor nodes.  The code logic remains exactly
     the same as before, however bigger changes in heat values are now
     propagated faster.}
\end{topfig}
\normalsize

Fig.~\ref{results:ht} presents the scalability results for the uncoordinated
and coordinated version. The dataset used is a square grid with an inner square
with high heat nodes. With 1 thread, there's a 33\% reduction in run time, while
for 16 threads, there's a 18\% reduction.

To further improve locality, we split the second rule to
not send small $\delta$ values if the target node is in another
thread (Fig.~\ref{code:ht_better}). The \textbf{Local-Only} line in
Fig.~\ref{results:ht} shows that improving locality by disabling small updates
improves the execution time of HT. XXX update figure

\begin{topfig}
\scriptsize\begin{Verbatim}[numbers=left,commandchars=\\\[\]]
new-heat(A, New, Old)
fabs(New - Old) <= epsilon
\underline[cpu-id(A, B, C)],
\underline[cpu-id(A, A, C)]
   -o {B, W | !edge(A, B, W) |
         new-neighbor-heat(B, A, New)}.

new-heat(A, New, Old)
fabs(New - Old) <= epsilon,
\underline[cpu-id(A, B, C1)],
\underline[cpu-id(A, A, C2)],
\underline[C1 <> C2]
   -o 1. // nothing is derived
\end{Verbatim}
  \scap{code:ht_better}{To improve locality, we added a third rule to not send small $\delta$
     values if the neighbor is in another thread.}
\end{topfig}
\normalsize

\begin{topfig}
   \begin{center}
      \subfloat[]{\includegraphics[width=5.5cm]{results/new-heat-transfer-80}}
   \end{center}
   \scap{results:ht}{Experimental results for the HT program. The coordinated version
      is, on average, 30\% faster than the regular version although it has a
      slightly worse scalability due to a reduction in work available.}
\end{topfig}