### 2014-11-12 related.tex
Update related work.

\paragraph{Declarative Languages}
Declarative programming paradigms such as \emph{logic} and \emph{functional}
programming have been extensively exploited for their implicit parallelism.
In logic languages such as Prolog, researchers took advantage of the
non-determinism of proof-search to evaluate subgoals in parallel with models
such as \emph{or-parallelism}~\cite{ali-86} and
\emph{and-parallelism}~\cite{Shen-92}.  In functional languages, the stateless
nature of computation allows multiple expressions to evaluate safely in
parallel. This has been explored in several languages such as
NESL~\cite{Blelloch:1996:PPA:227234.227246}, Id~\cite{Nikhil93anoverview}, and
more recently Data Parallel Haskell~\cite{nepal}.
NESL often got good performance, but was limited to a particular
application space. We also limit our application space, but also give the
programmer a declarative way to coordinate computation to improve the
performance of the program.

\paragraph{Data Centric Languages}
Recently, there has been an increasing interest in declarative data-centric
languages. MapReduce~\cite{Dean:2008:MSD:1327452.1327492}, for instance, is a
popular programming model that is optimized for large clusters.
Intrinsic to its popularity is the simplicity of its scheduling and data sharing
model.  In order to facilitate the writing of programs over large datasets,
SQL-like languages such as PigLatin~\cite{Olston:2008:PLN:1376616.1376726} have
been developed.  PigLatin builds on top of MapReduce and allows the programmer
to write complex data-flow graphs, raising the abstraction and easy of
programmability of MapReduce programs.  An alternative to PigLatin/MapReduce is
Dryad~\cite{Isard:2007:DDD:1272996.1273005} that allows programmers to design
arbitrary computation patterns using the DAG abstraction. It combines
computational vertices with communication channels (edges) that are
automatically scheduled to run on multiple computers/cores.

GraphLab~\cite{GraphLab2010} is a C++ framework for developing graph-based parallel machine
learning algorithms. GraphLab allows nodes to have read/write access to
different scopes through different concurrent access models in order to balance
performance and data consistency.  GraphLab also provides different schedulers
that dictate the order in which node's are computed, allowing the programmer to
optimize the program. Later in this paper, we will show how one GraphLab
scheduler can be implemented in CLM through the use of coordination facts.
Another, more restrictive, graph-based system is
Pregel~\cite{Malewicz:2010:PSL:1807167.1807184}, where graph algorithms must be
executed as a sequence of iterations of computation and message passing.
Both those systems provide excellent performance, however they are not
declarative and have a steep learning curve.

\paragraph{Coordination Languages}
Many programming languages follow the so-called \emph{coordination
paradigm}~\cite{Papadopoulos98coordinationmodels}, a form of distributed
programming that divides execution in two parts: \emph{computation}, where the actual
computation is performed, and \emph{coordination}, which deals with
communication and cooperation between processing units. This paradigm attempts
to clearly distinguish between these two parts by providing abstractions for
coordination in an attempt to provide architecture and system-independent forms
of communication.

Linda~\cite{linda} is arguably the most famous coordination model. Linda
implements a data-driven coordination model and features a \emph{tuple space}
that can be manipulated using the following coordination directives:
\texttt{out(t)} to write a tuple \texttt{t} into the tuple space; \texttt{in(t)}
to read a tuple using the template \texttt{t}; \texttt{rd(t)} to retrieve a copy
of the tuple \texttt{t} from the tuple space; and \texttt{eval(p)} to add a
process \texttt{p} in the tuple space and execute it in parallel.  Linda is be
implemented on top of many popular languages by simply creating a communication
and storage mechanism for the tuple space and then adding the directives as a
language library.

Another early coordination language is Delirium~\cite{Delirium}. Unlike Linda
which is embedded into another language, Delirium actually embeds operators
written in other languages inside the Delirium language.  The advantages of
Delirium are improved abstraction and easier debugging because sequential
operators are isolated from the coordination language.

Linda and Delirium are limited in the sense that the programmer can only
coordinate the scheduling of processing units, while the placement of data is
left to the implementation. But even more importantly, CLM differs from those
languages because coordination acts on data instead of processing units.
The abstraction is then raised by considering data and algorithmic aspects of
the program instead of focusing how processing units are used.
Furthermore, the CLM language is both a coordination language and a computation
language and there is no distinction between the two components.

\paragraph{Forward-Chaining Logic Programming}

Forward-chaining logic programming came into its own with the Datalog
language~\cite{Ullman:1990:PDK:533142}.  Traditionally used in deductive
databases, it is now being increasingly used in different fields such as
distributed networking~\cite{Loo-condie-garofalakis-p2}, sensor
nets~\cite{Chu:2007:DID:1322263.1322281} and cloud computing~\cite{alvaro:boom}.
The LM language was inspired by Meld~\cite{ashley-rollman-iclp09}, a
Datalog-like language for programming distributed ensembles of modular robots.
Meld introduced the idea of sensing and action facts in order to sense and act
on the outside world, respectively. LCM uses sensing facts to sense the
underlying system that the program is running on and action facts act on that
system.