### 2014-11-12 sssp.tex
spacing

Here we add coordination to the SSSP program described in \sectref{sect:ssspex}.
The coordinated version of the
SSSP~(Fig.~\ref{code:shortest_path_program_coord}) uses the coordination fact
\texttt{set-priority} (line 14). We also use a global program directive to order
priorities in ascending order (line 5).

When run with one thread, the algorithm behaves like
Dijkstra's shortest path algorithm~\cite{Dijkstra}. When using multiple
threads, each thread will pick the smallest distance from their subset of nodes.
While this does not yield the optimal program with relation to 1 thread, it
allows for parallel execution and locally avoids unnecessary work. The result
scales well and it is close to Dijkstra's algorithm.

\begin{topfig}
\scriptsize\begin{Verbatim}[numbers=left,commandchars=\\\{\}]
type route edge(node, node, int).
type linear shortest(node, int, list int).
type linear relax(node, int, list int).

\underline{priority @order asc}.

shortest(A, +00, []).
relax(@1, 0, [@1]).

shortest(A, D1, P1), D1 > D2, relax(A, D2, P2)
   -o shortest(A, D2, P2),
      \{B, W | !edge(A, B, W) |
         relax(B, D2 + W, P2 ++ [B]),
         \underline{set-priority(B, float(D2 + W))}\}.

shortest(A, D1, P1), D1 <= D2, relax(A, D2, P2)
   -o shortest(A, D1, P1).
\end{Verbatim}
  \scap{code:shortest_path_program_coord}{Shortest Path Program.}
\end{topfig}
\normalsize

The most interesting property of the SSSP program presented in
Fig.~\ref{code:shortest_path_program_coord} is that it remains provably correct,
although it applies rules using smarter ordering. The derivation of
\texttt{set-priority} does not change the behavior of the logical rules and the
code remains declarative.

Fig.~\ref{results:sssp_uspowergrid} shows experimental results when the
program computes the SSSP for 20\% of the nodes of the graph.
There are some situations where unnecessary facts are propagated
because although the shortest distance is selected locally, sub-optimal distances may be
propagated because many SSSP distances are computed at the same time.
However, we see a reduction of over 50\% in the number of
derived facts when using the coordinated version \textbf{C} over the regular
version \textbf{R}. The results also show that the coordinated
version using 16 threads is 1.3 times faster than the regular version using 16
threads while still showing good scaling.

\begin{topfig}
\vspace*{-5ex}
   \begin{center}
      \subfloat[]{ \begin{tabular}[b]{ | c | c | c | }
         \hline                       
         \textbf{\# T} & \textbf{R} & \textbf{C} \\ \hline \hline
         1 & 333K & 206K \\
         2 & 300K & 210K \\
         4 & 316K & 208K \\
         8 & 328K & 211K \\
         16 & 343K & 212K \\
         \hline  
         \end{tabular}
         \normalsize
      }
      \subfloat[]{\includegraphics[width=5cm]{results/shortest-uspowergrid.png}}
   \end{center}
   \scap{results:sssp_uspowergrid}{Results for SSSP run on the 
   US power grid network.  On the left we show the number of facts
   derived for the regular \textbf{R} and coordinated
   version \textbf{C} using a variable number of threads \textbf{\#
   T}. On the right, we show the scalability of the regular and
   coordinated version. The speedup values are computed using the
   execution time for 1 thread.}
\vspace*{-3ex plus 0pt minus 1ex}
\end{topfig}