### 2014-11-12 abstract.tex
Add facts derived

Declarative programming has been hailed as a promising approach to parallel
programming since it hides the implementation details of parallelism away from the
programmer. However, its advantage has also been its downfall
as it leaves the programmer with no straightforward way
to optimize programs for performance.
In this paper, we introduce Coordinated Linear Meld, a concurrent forward-chaining linear logic programming
language, with a declarative way to coordinate the execution of the program
allowing the programmer to change
computation scheduling and data layout. Our approach allows the programmer to
write declarative parallel programs and then optionally use coordination 
to fine tune programs, while keeping the program fully declarative and easy to
reason about.