### 2014-11-13 application.tex
Several corrections. Reduce size of texttt.

To better understand how coordination facts are used, we next present some programs that
take advantage of them. In our experimental setup, we used a machine with
four AMD Six-Core Opteron TM 8425 HE (2100 MHz) chips (24 cores) and 64 GB of
DDR-2 667MHz (16x4GB) RAM, running GNU/Linux (kernel 3.15.10-201 64 bits).
We compiled our virtual machine using GCC 4.8.3 (g++) with the flags
\mytt{-O3 -std=c++11 -fno-rtti -march=x86-64}~\footnote{Implementation,
   example programs and program proofs available in \url{http://github.com/.../meld}}.

\subsection{Single Source Shortest Path}
\input{sssp}

\subsection{MiniMax}
\input{minimax}

\subsection{Heat Transfer}
\input{heat-transfer}

\subsection{Splash Belief Propagation}
\input{splash-bp}

\subsection{N Queens}
\input{nqueens}