### 2014-12-17 rebuttal.txt
final formatting

We greatly appreciate the reviews of our paper.  We clearly sparked
interest, but need to improve the related work and experimental
results sections.  We address both these issues below.

Each reviewer asked for more (but different) information which would
require us to remove something in the paper.  We will happily
take advice from the committee about what we should remove to make
room for something else.

Our primary goal with this paper was to show the expressive power of
making coordination a first class programming construct and chose to
include five examples to give readers a flavor of the power of CLM.
This approach highlights the conciseness and expressiveness of CLM as
well as the readability of a high-level declarative approach to
fine-tuning the performance of a program.

Related Work:

The reviews pointed out numerous systems that we did not cite.  Please
note that CLM stands alone in
- making coordination (both scheduling and partitioning) a first-class
  programming construct.
- supporting data-driven dynamic (at run-time) coordination,
  particularly for irregular data structures.

We do not claim novelty in the mechanisms, but rather in how they are
integrated into a declarative language - as a first class entity.  We
will do a better job of showing where the ideas come from.

In reference to works cited by the reviewers:

- Halide supports distinct coordination, but only as "pragmas" in the
  domain of image processing, i.e., regular data-structures.

- Galois supports a rich set of schedulers through customizable
  containers (see: Nguyen&Pingali), but it requires extra-language
  mechanisms to do things such as determining the location of data.

- Elixir's uses a pre-defined set of schedulers and limits
  coordination to scheduling.

- Grace supports scheduling through priorities.  (It is very much like GraphLab.)

- Ligra, also oriented towards graph processing, is limited to shared
  memory and only handles partitioning.

- Cilk, OpenMP and TBB all provide low-level imperative approaches for
  coordination.

We will expand the related work section, but stress the unique nature
of CLM which results in clean concise coordination, e.g., we can write
the Splash scheduler used for Splash BP less than 60 lines (versus 500
in GraphLab).  It is hard (or impossible) to write in all the above
languages.

Experiments:

Mea Culpa on not explaining the superlinear speed-up.  We didn't
implement the randomized scheduler needed for the best sequential
implementation.  We will address this in the final version.

We did compare CLM to an optimized C++ implementation of 2 programs
(HT and SBP): The CLM implementations are 1.8X-3X slower, but scale
equally well.  Our focus was on expressiveness and scalability, not
raw performance. (CLM is currently interpreted by an
unoptimized virtual machine. It could even be compiled.)  From an
expressiveness point of view, the GraphLab scheduler is 500 lines of
customized library code.  CLM's version is less than 1 page!  If
requested, we can also include a third-party solution to SSSP.  As
well as including several different synthetic graphs to test the
scalability over different data sets.

We can include comparisons to sequential solutions:
- interpreted solutions (Python which is 2x-10x slower than CLM)
- compiled solutions in C (which is 2-10x faster depending on the
  program)

Other Stuff (And addressing some of the smaller, but interesting
points taken up by the reviewers):

-> reviewer A:
- Regarding the benchmark sizes, BP uses 160,000 nodes and MiniMax uses
   3,600,000 nodes. SSSP computes around 5,000,000 shortest paths.  In
   fact, we have found that the larger the data set the better the
   scaling.  We can increase the sizes by 10x if the reviews prefer.
- CLM's busy-wait locks are ticket locks that use atomic
   instructions. We have scalable locks but they need more testing.
- CLM can do synchronous or asynchronous algorithms, depending on the
   needs of the programmer.  We have implemented map-reduce style
   algorithms, asynchronous algorithms, and algorithms with mutable
   graphs.
- get-priority and set-priority are not sufficient.  The temporary
  priority is reset each time the node is processed, which comes in
  quite handy in the splash scheduler, for example.
- The arguments to be indexed and such as described in another paper.
  Happy to forward to committee (or provide the citation if needed.)
- There was a typo in the paper.  We actually do a vertical split, not
  a horizontal split when we do work stealing.
- We do reference counting for garbage counting, but no locks are used
  and the threads do it individually and do not need to stop and
  collect data.
- The implementation of SSSP devolves to Dijkstra's algorithm on 1
  thread, on multiple threads there are times when a non-optimal node
  in the graph is chosen, but overall it approaches the lower bounds
  of Dijkstra's algorithm while allowing for significantly more
  parallelism.

-> reviewer B:
- Program proofs: we did not include them due to lack of space.
   However, we have proven interesting properties of our programs, e.g.,
   correctness, termination, dead-lock freedom, liveness, etc.  We can
   remove the references to program proofs in the final version. (that's
         another paper)
- With regards to the poorly constructed sentence at the end of section
   6.3, there are no additional errors.  The values computed all meet the
   error bound, but are different for different schedulers.
- We agree that a measure of rule complexity would be a great thing to
   have.  Any suggestions?  Obviously lines of code is not enough.  The
   "gut" intuition here is that CLM is easy to read and doesn't require
   many lines of code.  A better metric would be great.
- The difference between Coordinated/Coordinated and Coordinated/Regular:
   - Coordinated/Regular: speedup of the coordinated program on
      n-processors over the unannotated (e.g., regular program) on 1
      processor.
   - Coordinated/Coordinated: speedup of the coordinated program on
      n-processors over the coordinated program on 1 processor.

-> reviewer A and C:
- The scalability results were not presented up to 24 threads due to an
error in our presentation, that info will be in the final paper.
(Just so you know, the graphs continue to look the same up through 24
threads.)

Finally, we note that CLM is noteworthy since it is not restricted to
shared memory systems and can be easily extended to distributed
systems. That simply cannot be said for most of the other systems.
The use of linear logic, a sound logical framework, has allowed us to
easily prove properties about programs using coordination.

We thank the reviewers for the time and expertise they have invested
in these reviews.  We will of course avail ourselves of all the
comments on presentation, grammar, etc.  Again, many thanks.

LocalWords:  Halide pragmas customizable Pingali pre GraphLab Ligra
LocalWords:  Cilk OpenMP TBB Mea Culpa superlinear SBP scalability
LocalWords:  unoptimized CLM's SSSP MiniMax scalable liveness
LocalWords:  unannotated