### 2015-01-20 rebuttal2.txt
Update.

Coordination is a highly important programming language design issue, specially
when concerning declarative programming languages and parallelism. Our effort
is not a small increment because it makes dynamic coordination a first class
entity of a language and semantically equivalent to computation, allowing
the programmer to write arbitrary coordination algorithms. This is a novel design
that fits well under the scope of PLDI. Unfortunately there has been some mismatch between
this important contribution and the reviewer's perspective on the material.

We wrote the paper from a programming language perspective and thus some
systems were unfarily neglected from the related work.
However, those systems are not programming languages on their own right and thus
do not solve the programming language design issue. While we think they
solve many important implementation issues, the programming language
aspect remains unsolved. CLM is an elegant approach to this challenge.

We compare the contributions of those other systems:
- Halide supports distinct coordination, but only as "pragmas" in the domain of
image processing, i.e., regular data-structures.
- Galois supports a rich set of schedulers through customizable containers
(see: Nguyen&Pingali), but it requires extra-language mechanisms to do things
such as determining the location of data.
- Elixir's uses a pre-defined set of schedulers and limits coordination to
   scheduling.
- Grace supports scheduling through priorities. (It is very much like GraphLab)
- Ligra, also oriented towards graph processing, only handles partitioning.
- Cilk, OpenMP and TBB all provide low-level imperative approaches for
coordination.

CLM distinguishes itself by integrating both scheduling and partitioning into a
declarative programming language as a first class entity. In particular CLM
stands alone in:
- supporting data-driven dynamic (at run-time) coordination, particularly for
   irregular data structures.
- actually being a declarative programming language and not a runtime system,
   solving both the implementation and design issues.
- being able to specify complex coordination programs such as SBP in less than
   60 lines (versus 500 in GraphLab).  It is hard (or impossible) to write it in
   all the above languages.

Each reviewer asked for more (but different) information which would require us
to remove something in the paper.  We could have included more experiments and
data sets but instead we decided to include five programs that give readers a
flavor of the power of CLM. This approach highlights the conciseness and
expressiveness of CLM as well as the readability of a high-level declarative
approach to fine-tune the performance of programs. We remind our reviewers that
our system is competitive (2-3x) against GraphLab and performs reasonabily well
against C++ (2-10x slower).

% include other_stuff.txt