### 2015-01-20 rebuttal2.txt
Update rebuttal.

Coordination is a highly important programming language design issue,
specially when concerning declarative programming languages and
parallelism. Our contribution makes dynamic coordination a first class
entity of a declarative language and semantically equivalent to
computation. Like reviewer B mentioned, this orthogonal design for
writing coordinated algorithms will be of interest to many in the PLDI
community. Unfortunately, there has been some misunderstanding between
this important contribution and the other reviewer's perspective on
the material.

Each reviewer asked for more (but different) information which would
require us to remove something in the paper. We could have included
more experiments and data sets but instead we decided to include five
programs that give readers a flavor of the power of CLM. This approach
highlights the conciseness and expressiveness of CLM as well as the
readability of a high-level declarative approach to fine-tune the
performance of programs. Our results show that CLM is competitive
(2-3x) against GraphLab and performs reasonabily well against C++
(2-10x slower).

We wrote the paper from a programming language perspective and thus
some systems were unfarily neglected from the related work. However,
those systems are not programming languages on their own right and
thus do not solve the programming language design issue. While they
solve many important implementation issues, the programming language
aspect remains unsolved. CLM is an elegant approach to this challenge.

We briefly compare the contributions of those other systems:
- Halide supports distinct coordination, but only as "pragmas" in the
  domain of image processing, i.e., regular data-structures.
- Galois supports a rich set of schedulers through customizable
  containers (see: Nguyen&Pingali), but it requires extra-language
  mechanisms to do things such as determining the location of data.
- Elixir's uses a pre-defined set of schedulers and limits
  coordination to scheduling.
- Grace supports scheduling through priorities (it is very much like
  GraphLab).
- Ligra, also oriented towards graph processing, only handles
  partitioning.
- Cilk, OpenMP and TBB all provide low-level imperative approaches for
  coordination.

CLM distinguishes itself by integrating both scheduling and
partitioning into a declarative programming language as a first class
entity. In particular CLM stands alone in:
- supporting data-driven dynamic (at run-time) coordination,
  particularly for irregular data structures.
- actually being a declarative programming language and not a runtime
  system, solving both the implementation and design issues.
- being able to specify complex coordination programs such as SBP in
  less than 60 lines (versus 500 in GraphLab). It is hard (or
  impossible) to write it in all the above languages.

% include other_stuff.txt